---
title: "Quality control and downstream analysis"
author:
  - "trnguyen@ukaachen.de"
date: "Last update on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float:
      toc_collapsed: false
    toc_depth: 3
    number_sections: false
    theme: lumen
---
  
```{css zoom-lib-src, echo = FALSE}
script src = "https://ajax.googleapis.com/ajax/libs/jquery/3.4.1/jquery.min.js"
```

```{js zoom-jquery, echo = FALSE}
$(document).ready(function() {
  $('body').prepend('<div class=\"zoomDiv\"><img src=\"\" class=\"zoomImg\"></div>');
  // onClick function for all plots (img's)
    $('img:not(.zoomImg)').click(function() {
      $('.zoomImg').attr('src', $(this).attr('src')).css({width: '100%'});
      $('.zoomDiv').css({opacity: '1', width: 'auto', border: '1px solid white', borderRadius: '5px', position: 'fixed', top: '50%', left: '50%', marginRight: '-50%', transform: 'translate(-50%, -50%)', boxShadow: '0px 0px 50px #888888', zIndex: '50', overflow: 'auto', maxHeight: '100%'});
    });
    // onClick function for zoomImg
    $('img.zoomImg').click(function() {
      $('.zoomDiv').css({opacity: '0', width: '0%'}); 
    });
  });
```


```{r preparation, echo=FALSE, results='hide', include=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, , message=FALSE}
# _____INPUT THE PREPROCESSED SEURAT OBJECT FROM THE PIPELINE__________
# gc()
# rm(list = ls())

set.seed(42)

path.to.pipeline.src <- "/home/hieunguyen/CRC1382/src/single_cell_pipeline_23052023/scRNA_GEX_pipeline"

source(file.path(path.to.pipeline.src, "processes_src", "import_libraries.R"))
source(file.path(path.to.pipeline.src, "processes_src", "helper_functions.R"))

####-----------------------------------------------------------------------#####
#### INPUT ARGUMENTS FOR THE HTML REPORTS
####-----------------------------------------------------------------------#####
# analysis.round <- params$analysis.round
# pct_mito <- params$pct_mito
analysis.round <- "1st_round"
pct_mito <- 25
####-----------------------------------------------------------------------#####

path.to.pipeline.output <- "/home/hieunguyen/CRC1382/src/single_cell_pipeline_23052023/OUTPUT"

analysis.round <- "1st_round"
pct.mito <- 10 
PROJECT <- "test_pbmc10k"

path.to.seurat.object <- file.path(path.to.pipeline.output, analysis.round, 
                                   sprintf("pct_mito_%s_1", pct.mito),
                                   "s8a_output",
                                   sprintf("%s.output.s8a.rds", PROJECT))

path.to.output <- file.path("/home/hieunguyen/CRC1382/src/single_cell_pipeline_23052023/OUTPUT", "data_analysis")
path.to.01.output <- file.path(path.to.output, "01_output")
dir.create(path.to.01.output, showWarnings = FALSE, recursive = TRUE)

s.obj <- readRDS(path.to.seurat.object)
```

# Preprocessing and QC
## Raw data Quality control  {.tabset}
```{r, results = 'asis', echo = FALSE}
for (plot.name in names(s.obj@misc$all.QC)){
  
  cat('###',plot.name,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  print(s.obj@misc$all.QC[plot.name])
  cat(' \n \n')
  
}

```


## Ambient RNA background correction 

### DecontX clusters {.tabset}
```{r, echo = FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, , results='asis'}

for (plot.name in names(s.obj@misc$ambient.cluster.RNA.plot)){
  
  cat('####',plot.name,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  print(s.obj@misc$ambient.cluster.RNA.plot[plot.name])
  cat(' \n \n')
}
```

### Contamination level in each sample

```{r}
ggplot(s.obj@meta.data, aes(x=name, y=AmbientRNA)) + 
  geom_boxplot()

```


### Ambient RNA contamination level {.tabset}

```{r, results = 'asis', echo = FALSE}
for (plot.name in names(s.obj@misc$ambient.contamination.plot)){
  
  cat('####',plot.name,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  print(s.obj@misc$ambient.contamination.plot[plot.name])
  cat(' \n \n')
}
```


## Descriptive statistics and filtering threshold {.tabset}
This section is devoted to the descriptive statistics of the following varialbes: `nFeature_RNA, nCount_RNA, percent.mt, percent.ribo`. 

```{r, results = 'asis', echo = FALSE}
for (plot.item in c("nFeature_RNA", "nCount_RNA", "percent.mt", "percent.ribo")){
  boxplot <- s.obj@meta.data %>% ggplot(aes_string(x = "name", y = plot.item)) +
    geom_boxplot(outlier.colour="black", outlier.shape=16,
                 outlier.size=2, notch=FALSE) +
    ggtitle(sprintf("Boxplot: Distribution of %s in each dataset", plot.item))
  cat('###', plot.item,'{.unlisted .unnumbered} \n')
  
  ##### 
  # plots or tables that we want to show in tabs
  #####
  
  print(boxplot)
  
  cat(' \n \n')
}
```

## Descriptive statistics + UMAP {.tabset}

### % Mitochondrial
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, }
FeaturePlot(object = s.obj, reduction = "RNA_UMAP", feature = "percent.mt", label.size = 6, label = TRUE, label.color = "red")
```

### % Ribosome
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, }
FeaturePlot(object = s.obj, reduction = "RNA_UMAP", feature = "percent.ribo", label = TRUE, label.color = "red", label.size = 6)
```

### % nCount RNA
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, }
FeaturePlot(object = s.obj, reduction = "RNA_UMAP", feature = "nCount_RNA", label = TRUE, label.color = "red", label.size = 6)
```

### % nGenes 
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, }
FeaturePlot(object = s.obj, reduction = "RNA_UMAP", feature = "nFeature_RNA", label = TRUE, label.color = "red" , label.size = 6)
```

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, }
DimPlot(s.obj, reduction = "RNA_UMAP", split.by = "Phase")

```

# UMAP and clustering

### One sample per plot {.tabset}
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, }
library(scales)
colors <- hue_pal()(3)
all.samples <- unique(s.obj$name)
for (i in seq(1, length(unique(s.obj$name)))){
  sample.id <- all.samples[[i]]
  cat(sprintf("#### Sample: %s \n", sample.id))
  c <- colors[[i]]
  sample.cells <- colnames(subset(s.obj, name == sample.id))
  names(sample.cells) <- sample.id
  p <- DimPlot(object = s.obj, reduction = "RNA_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, cells.highlight = sample.cells, cols.highlight = c, pt.size = 0.5)  + theme(legend.position = "none") + ggtitle(sample.id)
  print(p)
  cat("\n \n")
}

```

## After integration
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, }
DimPlot(object = s.obj, reduction = "RNA_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, group.by = "name")
```

### One sample per plot {.tabset}
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, }
library(scales)
colors <- hue_pal()(3)
all.samples <- unique(s.obj$name)
for (i in seq(1, length(unique(s.obj$name)))){
  sample.id <- all.samples[[i]]
  cat(sprintf("#### Sample: %s \n", sample.id))
  c <- colors[[i]]
  sample.cells <- colnames(subset(s.obj, name == sample.id))
  p <- DimPlot(object = s.obj, reduction = "RNA_UMAP", label = TRUE, label.box = TRUE, repel = TRUE, cells.highlight = sample.cells, cols.highlight = c, pt.size = 0.5) + theme(legend.position = "none") + ggtitle(sample.id)
  print(p)
  cat("\n \n")
}

```


# Clusterings and cluster markers identification
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, }
DimPlot(object = s.obj, reduction = "RNA_UMAP", label = TRUE, label.box = TRUE)
```

## Cluster markers {.tabset}
```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, }
if (file.exists(file.path(path.to.01.output, "DE_cluster_marker_genes.rds")) == FALSE){
  diff.cluster.markers <- FindAllMarkers(object = s.obj, assay = "RNA", test.use = "wilcox",)
  diff.cluster.markers <- subset(diff.cluster.markers, diff.cluster.markers$p_val_adj < 0.05 & diff.cluster.markers$avg_log2FC > 0)
  saveRDS(diff.cluster.markers, file.path(path.to.01.output, "DE_cluster_marker_genes.rds"))
} else {
  diff.cluster.markers <- readRDS(file.path(path.to.01.output, "DE_cluster_marker_genes.rds"))
}

dir.create(file.path(path.to.01.output, "cluster_marker_genes_tables"), showWarnings = FALSE)

for (cluster.id in unique(diff.cluster.markers$cluster)){
  tmp_table <- subset(diff.cluster.markers, diff.cluster.markers$cluster == cluster.id) %>%
    rowwise %>%
    mutate_if(is.numeric, round, 8) %>%
    arrange(desc(avg_log2FC))
  
  writexl::write_xlsx(x = tmp_table, path = file.path(path.to.01.output, "cluster_marker_genes_tables", sprintf("cluster_%s.xlsx", cluster.id)))
}
```


## All cluster marker genes table {.tabset}
```{r results='hide', echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8}
diff.cluster.markers %>% create_dt()
```


```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8}
for (cluster.id in unique(diff.cluster.markers$cluster)){
  tmp.table <- subset(diff.cluster.markers, diff.cluster.markers$cluster == cluster.id)
  tmp.table <- tmp.table %>% mutate_if(is.numeric, round, 6)
  row.names(tmp.table) <- NULL
  cat(paste("\n\n### Cluster: ", cluster.id, "##\n"))
  print( htmltools::tagList(datatable(tmp.table, extensions = 'Buttons',
                filter = "top",
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All")),
                               columnDefs = list(list(
                                 targets = "_all",
                                 render = JS(
                                   "function(data, type, row, meta) {",
                                   "return type === 'display' && data != null && data.length > 100 ?",
                                   "'<span title=\"' + data + '\">' + data.substr(0, 100) + '...</span>' : data;",
                                   "}")
                               ))
                ))))
  cat("\n \n")
}
```

## Top-16 cluster markers feature plot {.tabset}

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, , fig.width=12, fig.height=8}
for (cluster.id in unique(diff.cluster.markers$cluster)){
  DefaultAssay(s.obj) <- "RNA"
  tmp <- subset(diff.cluster.markers, diff.cluster.markers$cluster == cluster.id & diff.cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  arrange(desc(avg_log2FC)) %>% head(16)
  cat(sprintf("### Cluster %s \n", cluster.id))
  p <- FeaturePlot(object = s.obj, reduction = "RNA_UMAP", features = tmp$gene, ncol = 4, ) 
  print(p)
  cat("\n \n")
}
```

## Top-16 cluster markers dot plot {.tabset}

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, , fig.width=12, fig.height=8}
for (cluster.id in unique(diff.cluster.markers$cluster)){
  DefaultAssay(s.obj) <- "RNA"
  tmp <- subset(diff.cluster.markers, diff.cluster.markers$cluster == cluster.id & diff.cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  arrange(desc(avg_log2FC)) %>% head(16)
  cat(sprintf("### Cluster %s \n", cluster.id))
  p <- DotPlot(object = s.obj, features = tmp$gene) 
  print(p)
  cat("\n \n")
}
```

## Top-16 cluster markers Violin plot {.tabset}

```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, , fig.width=12, fig.height=8}
for (cluster.id in unique(diff.cluster.markers$cluster)){
  DefaultAssay(s.obj) <- "RNA"
  tmp <- subset(diff.cluster.markers, diff.cluster.markers$cluster == cluster.id & diff.cluster.markers$p_val_adj < 0.05) %>% 
  rowwise %>% 
  mutate_if(is.numeric, round, 6) %>% 
  arrange(desc(avg_log2FC)) %>% head(16)
  cat(sprintf("### Cluster %s \n", cluster.id))
  p <- VlnPlot(object = s.obj, features = tmp$gene, pt.size = 0, ) 
  print(p)
  cat("\n \n")
}
```


```{r results='asis', echo=FALSE, message=FALSE, warning=FALSE, fig.width = 14, fig.height = 10, , fig.width=25, fig.height=15}
```



